---
title: "EDAV Final Data Prep"
author: Myles Ingram, Colin Payne-Rogers
output:
  html_document:
    fig_height: 5
    fig_width: 9
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  cache = TRUE,
  message = FALSE,
  warning = FALSE)
```

# Data transformation

The goal of this markdown file, as a draft, is to load raw project data from file and save that data in a form more conducive to future analyses. This format will change when draft chapters are combined into a final book.

The raw data for the project consists of two `.csv` files and one `.xls` file. Each `.csv` corresponds to digestive system cancer incidence and digestive system cancer mortality, respectively, for the years 2000 to 2017. The `.xls` file contains information on per capita consumption from the years 1970 to the present.

## Digestive Cancer Incidence

First, load the `.csv` data from file.

```{r, eval=TRUE}
incidence <- read.csv(
  file='resources/raw_data/digestive_cancer_incidence.csv',
  header=FALSE,
  stringsAsFactors=TRUE)
head(incidence)
```

Remove the summary rows, whose first column contains `2000-2017`.

```{r, eval=TRUE}
incidence <- incidence[!incidence$'"V1"' == "2000-2017",]
head(incidence)
```
