# Results

```{r}
library(ggplot2)
library(dplyr)
library(tidyverse)
library(plyr)
require(scales)
```

```{r, include=FALSE}
dairy <- read_csv("resources/processed_data/dairy.csv")
fats <- read_csv("resources/processed_data/fats.csv")
fruits <- read_csv("resources/processed_data/fruits.csv")
grains <- read_csv("resources/processed_data/grains.csv")
incidence <- read_csv("resources/processed_data/incidence.csv")
meateggsnuts <- read_csv("resources/processed_data/meateggsnuts.csv")
mortality <- read_csv("resources/processed_data/mortality.csv")
sugars <- read_csv("resources/processed_data/sugars.csv")
vegetables <- read_csv("resources/processed_data/vegetables.csv")
total_cals <- read_csv("resources/processed_data/total.csv")

tidy_calorie_data <- function(food_table) {
  df <- subset(gather(food_table, food_group, calories,    colnames(food_table), -Year), Year > 2000)
  return(df)
}
```

1. Line graph of mortality and incidence
2. Line graph of total caloric intake and caloric intake by major food group
3. Line graph of specific foods
4. Scatterplot of total caloric intake and caloric intake by major food group vs cancer mortality
5. Scatterplot of total caloric intake and caloric intake by major food group vs cancer incidence
6. Scatterplot of specific foods vs cancer incidence/mortality
6. Bargraph of R-squared values/correlation coefficients

```{r}
ggplot(data=incidence)+
  geom_line(aes(x=Year, y=V3, color=Sex)) +
  ggtitle("Digestive Cancer Incidence Rates")+
  ylab("Rate (per 100,000 people)")

ggplot(data=mortality)+
  geom_line(aes(x=Year, y=V3, color=Sex)) +
  ggtitle("Digestive Cancer Mortality Rates")+
  ylab("Rate (per 100,000 people)")

```

```{r}
tidy_total = tidy_calorie_data(total_cals[-c(9)])
tidy_all = tidy_calorie_data(total_cals[c(1,9)])

ggplot(data=tidy_total) +
  geom_line(aes(x=Year, y=calories, color=food_group)) +
  ggtitle("Per Capita Calorie Intake from 1970-2017")
ggplot(data=tidy_all) +
  geom_line(aes(x=Year, y=calories, color=food_group)) +
  ggtitle("Per Capita Calorie Intake from 1970-2017")

```

```{r}
tidy_sugar = tidy_calorie_data(sugars[-c(1)])
ggplot(data=tidy_sugar) +
  geom_line(aes(x=Year, y=calories, color=food_group)) +
  ggtitle("Per Capita Calorie Intake from 2000-2017")

```
```{r}
tidy_dairy = tidy_calorie_data(dairy[c(2:12)])
tidy_dairy_2 = tidy_calorie_data(dairy[c(2,18:30)])
tidy_dairy_3 = tidy_calorie_data(dairy[c(2,30:37)])
ggplot(data=tidy_dairy) +
  geom_line(aes(x=Year, y=calories, color=food_group)) +
  ggtitle("Per Capita Calorie Intake from 2000-2017")
ggplot(data=tidy_dairy_2) +
  geom_line(aes(x=Year, y=calories, color=food_group)) +
  ggtitle("Per Capita Calorie Intake from 2000-2017")
ggplot(data=tidy_dairy_3) +
  geom_line(aes(x=Year, y=calories, color=food_group)) +
  ggtitle("Per Capita Calorie Intake from 2000-2017")
```
```{r}
tidy_prot = tidy_calorie_data(meateggsnuts[c(2:15)])
tidy_prot_2 = tidy_calorie_data(meateggsnuts[c(2,15:26)])
ggplot(data=tidy_prot) +
  geom_line(aes(x=Year, y=calories, color=food_group)) +
  ggtitle("Per Capita Calorie Intake from 2000-2017")
ggplot(data=tidy_prot_2) +
  geom_line(aes(x=Year, y=calories, color=food_group)) +
  ggtitle("Per Capita Calorie Intake from 2000-2017")
```

#Scatter plots
```{r}
total_cals_inc = merge(total_cals, incidence, by="Year")
total_cals_inc$`Sugar and sweeteners (Added)`
ggplot(total_cals_inc) +
  geom_point(aes(`Sugar and sweeteners (Added)`, V3, color=Sex))

sugar_inc = merge(sugars, incidence, by="Year")
sugar_inc$`High Fructose Corn Syrup`
ggplot(sugar_inc) +
  geom_point(aes(`High Fructose Corn Syrup`, V3, color=Sex))

meat_inc = merge(meateggsnuts, incidence, by="Year")
ggplot(meat_inc) +
  geom_point(aes(Beef, V3, color=Sex))
```


