# Results

```{r}
library(ggplot2)
library(tidyverse)
library(grid)
require(gridExtra)
require(scales)
library(dplyr)
library(patchwork)
```

```{r include=FALSE}
dairy <- read_csv("resources/processed_data/dairy.csv")
fats <- read_csv("resources/processed_data/fats.csv")
fruits <- read_csv("resources/processed_data/fruits.csv")
grains <- read_csv("resources/processed_data/grains.csv")
incidence <- read_csv("resources/processed_data/incidence.csv")
meateggsnuts <- read_csv("resources/processed_data/meateggsnuts.csv")
mortality <- read_csv("resources/processed_data/mortality.csv")
sugars <- read_csv("resources/processed_data/sugars.csv")
vegetables <- read_csv("resources/processed_data/vegetables.csv")

Year = dairy$Year
`Meat, eggs, and nuts` <- rowSums(meateggsnuts[,c(2:length(meateggsnuts))], na.rm=TRUE)
Dairy <- rowSums(dairy[,c(2:length(dairy))], na.rm=TRUE)
Fruit <- rowSums(fruits[,c(2:length(fruits))], na.rm=TRUE)
Vegetables = rowSums(vegetables[,c(2:length(vegetables))], na.rm=TRUE)
`Flour and cereal products` <- rowSums(grains[,c(2:length(grains))], na.rm=TRUE)
`Added fats and oils and dairy fats` <- rowSums(fats[,c(2:length(fats))], na.rm = TRUE)
`Sugar and sweeteners (Added)` <- rowSums(sugars[,c(2:length(sugars))], na.rm=TRUE)

total_cals <- data.frame(Year, `Meat, eggs, and nuts`, Dairy, Fruit, Vegetables, `Flour and cereal products`, `Added fats and oils and dairy fats`, `Sugar and sweeteners (Added)`, check.names = FALSE)

# Large portions of the data are NA so removing the data from these years in totals
total_cals$`Added fats and oils and dairy fats`[42:48] = NA

total_cals$Total <- rowSums(total_cals[c(2:length(total_cals))])

tidy_calorie_data <- function(food_table) {
  df <- subset(gather(food_table, food_group, calories, colnames(food_table), -Year), Year > 1999)
  return(df)
}
```

We'll start our journey through the exciting world of digestive system cancer by looking at the trends in cancer incidence and mortality in recent years. Then, we'll break down (*digest*) trends in diet at the level of food group and top food sources. That'll give us a sense of each data set separately before we dive more deeply into possible relationships between diet and digestive cancer.

## Digestive System Cancer Trends

```{r}
# generates the cancer incidence plots
inc <-ggplot(data=incidence)+
  geom_line(aes(x=Year, y=Rate, color=Sex)) +
  geom_point(aes(x=Year, y=Rate, color=Sex))+
  scale_color_manual(values=c("black", "red", "blue"))+
  theme(legend.position = "none")+
  ggtitle("Incidence")+
  ylim(10, 150)+
  ylab("Rate (per 100,000 people)")

# generates the cancer mortality plots
mort <- ggplot(data=mortality)+
  geom_line(aes(x=Year, y=Rate, color=Sex)) +
  geom_point(aes(x=Year, y=Rate, color=Sex))+
  ggtitle("Mortality")+
  ylim(10,150)+
  ylab("Rate (per 100,000 people)")+
  scale_color_manual(values=c("black", "red", "blue"))

# fix axes
# takes the incidence and mortality plots and puts them on the same line with shared legend (1)
combined <- inc + mort & theme(legend.position = "bottom")
combined + plot_layout(guides = "collect") + plot_annotation("Digestive System Cancer Incidence and Mortality Rates From 2000 to 2017")
```

Both the digestive system incidence and mortality rates have been decreasing since 2000. Digestive cancer mortality contains a distinct, nonlinear decrease since 2014 or 2015. Men are at a greater risk of developing digestive cancer compared to women, but the trends of incidence and mortality for both genders mirror one another. Now that we have established the trends of digestive system cancer, let's explore dietary trends from 2000 to 2017.

## Dietary Trends

### Overall Consumption

```{r}
# sums all the columns except for the fats and oils column to get the total amount of calories consumed besides the fats and oils
total_cals$`Total (excluding Fats and Oils)` <- rowSums(total_cals[,c(2:6, 8)])

# creates two separate dfs: one for the food groups (tidy_total) and one for the totals (tidy_all)
tidy_total = tidy_calorie_data(total_cals[-c(9,10)])
tidy_all = tidy_calorie_data(total_cals[c(1,9:10)])

cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
# graph of the per capita total calorie intake and total calorie intake excluding fats and oils
ggplot(data=tidy_all) +
  geom_line(aes(x=Year, y=calories, color=food_group)) +
  geom_point(aes(x=Year, y=calories, color=food_group, shape=food_group))+
  ylab("Daily Per Capita Calories")+
  labs(color = "Calorie Totals", shape = "Calorie Totals")+
  ggtitle("Total Daily Per Capita Calorie Intake from 2000-2017")

# graph of the per capita calorie intake by major food groups
ggplot(data=tidy_total) +
  scale_fill_manual(values=cbPalette)+
  geom_line(aes(x=Year, y=calories, color=food_group)) +
  geom_point(aes(x=Year, y=calories, color=food_group, shape=food_group))+
  ylab("Daily Per Capita Calories")+
  labs(color = "Food Groups", shape = "Food Groups")+
  ggtitle("Daily Per Capita Calorie Intake by Major Food Group from 2000-2017")
```

These graphs show the trend of both the per capita total daily calorie intake and the calorie intake divided by major food groups. The most notable changes for the food groups were sugar and sweeteners, flour and cereal products, and meat, eggs, and nuts. Both sugar and sweeteners and flour and cereal products decreased from 2000 to 2017 while meat, eggs, and nuts increased from 2014-2017 after gradually decreasing before. Intake of fruits and vegetables has also slightly decreased over the years. In terms of calories, both total calories and total calories, excluding fats and oils have been decreasing from 2000-2017 as a whole. However, from 2011 to 2017 total calories excluding fats and oils has risen slightly. Calorie intake data for fats and oils from 2010 onward could not be found leading to the creation of the total calorie intake excluding fats and oils variable. Now that we know that calorie intake has been decreasing as a whole and within many major food groups, let's look at certain foods within each food group to discover more trends.

### Select Food

```{r}
# merges all the separate food group dataframes into one dataframe joined by year
allfoods <- merge(meateggsnuts, sugars, by="Year")
allfoods <- merge(allfoods, dairy, by="Year")
allfoods <- merge(allfoods, fats, by="Year")
allfoods <- merge(allfoods, fruits, by="Year")
allfoods <- merge(allfoods, grains, by="Year")
allfoods <- merge(allfoods, vegetables, by="Year")

# converts the dataframe into a tidy format and changes the calorie values into numeric
allfoods_tidy <- tidy_calorie_data(allfoods)
allfoods_tidy$calories <- as.numeric(allfoods_tidy$calories)
allfoods_na <- allfoods_tidy %>% dplyr::group_by(food_group) %>% dplyr::filter(!any(is.na(calories)))

# gets the percent difference from 2000 for each year for each food
allfoods_tidy <- allfoods_tidy %>% dplyr::group_by(food_group) %>%
  dplyr::mutate(perc_dif = round((calories-first(calories))/first(calories)*100, 4))

# makes a dataframe for the mean calorie intake for each food
meanfoods <- allfoods_tidy %>% dplyr::group_by(food_group) %>% dplyr::summarise(Mean = round(mean(calories, na.rm = TRUE), 2))

# takes the top 25 foods that have the highest mean calorie intake for analysis and gets rid of foods that have na values for certain years
topfoods <- head(meanfoods %>% arrange(desc(Mean)), 25)
meannafoods <- allfoods_na %>% dplyr::group_by(food_group) %>% dplyr::summarise(Mean = round(mean(calories, na.rm = TRUE), 2))
topnafoods <- head(meannafoods %>% arrange(desc(Mean)), 25)

# bar graph of the top 25 foods that contribute the most to total caloric intake
topnafoods %>%
  mutate(food_group=factor(food_group, levels=food_group)) %>%
  ggplot()+
    geom_bar(aes(x=food_group, y=Mean), stat="identity") +
    ylab("Mean Daily Per Capita Calories")+
    xlab("Foods")+
    ggtitle("Top 25 Calorie Intake Food Sources")+
    theme(axis.text.x = element_text(angle=60, hjust=1))

# splits the 25 foods into low medium and high to increase clarity of the line plot
topnafoods_tidy <- subset(allfoods_tidy, food_group %in% topnafoods$food_group)
topnafoods_tidy$calories_consumed <- cut(topnafoods_tidy$calories, breaks = c(0,50,100,500), label=c("10+", "50+", "100+"))

medhighfoods <- subset(topnafoods_tidy, calories_consumed %in% c("100+", "50+"))
lowfoods <- subset(topnafoods_tidy, calories_consumed %in% c("10+"))

# line plot of the top 25 foods
ggplot(data=medhighfoods) +
  geom_line(aes(x=Year, y=calories, color=food_group)) +
  geom_point(aes(x=Year, y=calories, color=food_group, shape=food_group))+
  ggtitle("50+ Calorie Intake Top 25 Food Sources from 2000-2017")+
  ylab("Mean Daily Per Capita Calories")+
  xlab("Years")+
  labs(color = "Food Groups", shape = "Food Groups")+
  facet_wrap(~calories_consumed) +
  theme(axis.text.x = element_text(angle=60, hjust=1))+
  theme(legend.position = "right", legend.text=element_text(size=5))+
  scale_shape_manual(values = 0:25)

ggplot(data=lowfoods) +
  geom_line(aes(x=Year, y=calories, color=food_group)) +
  geom_point(aes(x=Year, y=calories, color=food_group, shape=food_group))+
  ggtitle("10+ Calorie Intake Top 25 Food Sources from 2000-2017")+
  ylab("Mean Daily Per Capita Calories")+
  xlab("Years")+
  guides(fill=guide_legend(ncol=2))+
  labs(color = "Food Groups", shape = "Food Groups")+
  facet_wrap(~calories_consumed) +
  theme(axis.text.x = element_text(angle=60, hjust=1))+
  theme(legend.position = "bottom", legend.text=element_text(size=5))+
  scale_shape_manual(values = 0:25)


```

To determine how diet relates to digestive cancer incidence and mortality, we took the top 25 foods that contributed most to daily caloric intake and analyzed their changes from 2000 to 2017. These included several sources that have been related to digestive cancer, such as beef and high fructose corn syrup. Other foods that surprisingly (to us) cracked the top 25 were bananas, cane/beet sugar, and  frozen potatoes. The highest calorie contributor (by volume) was white and whole flour by a large margin. All of these top food sources changed dynamically from 2000 to 2017, with foods such as peanuts, chicken, and butter increasing and other foods such as beef, high fructose corn syrup, and ice cream decreasing. Now that we've set the stage, let's jump into the comparisons between cancer incidence/mortality and foods/total calorie intake.

## Diet and Digestive Cancer

The destination of our journey. Let's see how foods, food groups, and caloric intake relate to digestive cancer.

### Cancer Incidence
```{r}
# uses the combined sex since male and female cancer have the same trend
sub_inc <- read_csv("resources/processed_data/combined_incidence.csv")

# used to combine the incidence and major food group/total calories dataframes then tidy
total_cals_inc = merge(total_cals, sub_inc, by="Year")
canc_cals_tidy <- gather(total_cals_inc, food_group, calories, colnames(total_cals_inc)[!colnames(total_cals_inc) %in% c("Year", "Sex", "Population", "Cases")], -c(Year, Rate))
canc_cals_tidy <- canc_cals_tidy[complete.cases(canc_cals_tidy),]

# Splits the dataframe into the major food groups and the totals
canc_food_tidy <- subset(canc_cals_tidy, !food_group %in% c("Total", "Total (excluding Fats and Oils)"))
canc_total_tidy <- subset(canc_cals_tidy, food_group %in% c("Total", "Total (excluding Fats and Oils)"))

# Graph between the calories
ggplot(canc_food_tidy, aes(calories, Rate))+
    geom_point()+
    geom_smooth(method=lm, color="blue", fill="blue")+
    ylab("Rate (per 100,000 people)")+
    xlab("Daily Per Capita Calorie Intake")+
    ggtitle("Cancer Incidence vs Calorie Intake per Major Food Group")+
    facet_wrap(~food_group)
```

Even using a smoother to draw a line through the data, the correlations are unclear using the raw caloric intake values. We can see that both "sugar and sweeteners" and "flour and cereal products" correlate with digestive cancer incidence but the trends of the other groups are unclear. We'll use percent difference from 2000 per capita caloric intake from now on to show the trends more clearly.

```{r}
# Gets the percent difference from 2000
canc_cals_tidy <- canc_cals_tidy %>% dplyr::group_by(food_group) %>%
  dplyr::mutate(perc_dif = round((calories-first(calories))/first(calories)*100, 4))

# Creates a dataframe that has the correlation coefficient between the food group/total and cancer incidence and p_value for correlation
canc_cal_cor <- canc_cals_tidy %>% dplyr::group_by(food_group) %>% dplyr::summarise(Correlation = round(cor(perc_dif, Rate), 2))
temp1 <- canc_cals_tidy %>% dplyr::group_by(food_group) %>% dplyr::summarise(p_value = round(cor.test(perc_dif, Rate)$p.value, 4))
canc_cal_cor$p_value <- temp1$p_value

# Cancer Incidence vs total caloric intake and major food groups
ggplot(canc_cals_tidy, aes(perc_dif, Rate))+
    geom_point()+
    geom_smooth(method=lm, color="blue", fill="blue")+
    ylab("Rate (per 100,000 people)")+
    xlab("Percent Difference Per Capita Calorie Intake from 2000")+
    ggtitle("Cancer Incidence vs Calorie Intake per Major Food Group/Total Calorie Intake")+
    facet_wrap(~food_group)

# Correlation coefficients for above graph
canc_cal_cor %>%
  arrange(desc(Correlation)) %>%
  mutate(food_group=factor(food_group, levels=food_group)) %>%
  ggplot(aes(x=food_group, y=Correlation))+
    geom_bar(stat="identity") +
    geom_text(aes(label=p_value), size=2, color="red", position = position_stack(vjust = .5))+
    annotate("text", label = "Red text within bar graph is the p value associated with correlation", x = 5, y = -1, size = 2, colour = "red")+
    ylab("Correlation Coefficient")+
    xlab("Food Groups")+
    ggtitle("Correlation Coefficients of Food Groups/Total Caloric Intake vs Cancer Incidence")+
    theme(axis.text.x = element_text(angle=60, hjust=1))
```

We found that the decrease in total calorie intake correlated with the decrease in digestive cancer incidence. We also found that food groups such as "sweeteners and sugars" and "cereal/flour products" positively correlated with digestive cancer incidence. Surprisingly, we found that fruits caloric intake and vegetable caloric intake correlated positively with digestive cancer incidence. We think that since total caloric intake is associated with cancer incidence, the other food groups appear to be correlated as well.

```{r}
#scat mat with incidence mortality caloric intake and food groups
total_cals_inc_2 <- total_cals_inc %>% dplyr::rename(
                                        Sugar = `Sugar and sweeteners (Added)`,
                                        Fats = `Added fats and oils and dairy fats`,
                                        `Total (no fats)` = `Total (excluding Fats and Oils)`,
                                        Protein = `Meat, eggs, and nuts`,
                                        Grain = `Flour and cereal products`
                                        )

cal_scatmat <- total_cals_inc_2 %>% dplyr::select(Fruit, Vegetables, Sugar, Fats, Protein, Grain, 
                                                  `Total (no fats)`, Total, Rate)
plot(cal_scatmat, main="Scatterplot Matrix of Major Food Groups and Cancer Incidence" )
```

Looking closely at this scatter plot matrix, our hunch about other major food groups being correlated with total caloric intake was correct. All major food groups trend upwards when compared to both total caloric intake and total caloric intake excluding fats. That doesn't mean we shouldn't look at specific foods more closely. To a food-poor nation, high-calorie food sources are gold. To a food-rich nation, the culture or status associated with these foods (or, the markets, supply chains, and industries that make them so available) might drive the overall trends we're seeing.

```{r}
total_foods_inc = merge(topnafoods_tidy, distinct(canc_cals_tidy[c("Year", "Rate")]), by="Year")
food_inc_cor <- total_foods_inc %>% dplyr::group_by(food_group) %>% dplyr::summarise(Correlation = round(cor(perc_dif, Rate), 2))
temp <- total_foods_inc %>% dplyr::group_by(food_group) %>% dplyr::summarise(p_value = round(cor.test(perc_dif, Rate)$p.value, 4))
food_inc_cor$p_value <- temp$p_value

# Cancer incidence and top 25 food calories
ggplot(total_foods_inc, aes(perc_dif, Rate))+
    geom_point()+
    geom_smooth(method=lm, color="blue", fill="blue")+
    ylab("Rate (per 100,000 people)")+
    xlab("Percent Difference Per Capita Calorie Intake from 2000")+
    ggtitle("Cancer Incidence vs Calorie Intake per Top 25 Foods")+
    theme(axis.text.x = element_text(size=5.5, angle = 45))+
    facet_wrap(~food_group)


# Correlation coefficients for above graph
food_inc_cor %>%
  arrange(desc(Correlation)) %>%
  mutate(food_group=factor(food_group, levels=food_group)) %>%
  ggplot(aes(x=food_group, y=Correlation))+
    geom_bar(stat="identity") +
    geom_text(aes(label=p_value), size=2, color="red", position = position_stack(vjust = .5))+
    annotate("text", label = "Red text within bar graph is the p value associated with correlation", x = 5, y = -1, size = 2, colour = "red")+
    ylab("Correlation Coefficient")+
    xlab("Foods")+
    ggtitle("Correlation Coefficients of Top 25 Foods vs Cancer Incidence")+
    theme(axis.text.x = element_text(angle=60, hjust=1))

```


Decomposing the food groups into the top 25 foods by caloric intake yields more interesting results. As expected, high fructose corn syrup and beef are positively associated with digestive cancer incidence. Ice cream, potatoes, and orange juice trend positively with incidence, too. On the other hand, we found that butter, peanuts, cane/beet sugar, and mozzarella trended negatively with cancer incidence, meaning the higher the calorie intake of these foods, the lower the cancer incidence.

Each of these trends are worth investigating in their own right. Given the complexity of the problem, it isn't likely that these data are saying that butter, or peanuts, or mozzarella cheese are protective against digestive cancer. But they might represent clusters of diets or cultures with sets of behavior and/or genetics that are protective against digestive cancer. For the most part, we'll leave those extrapolations to the reader. For a couple of the most surprising results (namely, fruits and vegetables ... super foods!) we'll look more deeply into the data.

```{r}
# top vegetables
veg_inc = merge(vegetables, sub_inc, by="Year")

veg_inc_tidy <- gather(veg_inc, foods, calories, colnames(veg_inc)[!colnames(veg_inc) %in% c("Year", "Sex", "Population", "Cases")], -c(Year, Rate))
veg_inc_tidy <- veg_inc_tidy[complete.cases(veg_inc_tidy),]

# converts the dataframe into a tidy format and changes the calorie values into numeric
veg_tidy <- tidy_calorie_data(vegetables)
veg_tidy$calories <- as.numeric(veg_tidy$calories)
veg_tidy <- veg_tidy %>% dplyr::group_by(food_group) %>% dplyr::filter(!any(is.na(calories)))

# gets the percent difference from 2000 for each year for each food
veg_tidy <- veg_tidy %>% dplyr::group_by(food_group) %>%
  dplyr::mutate(perc_dif = round((calories-first(calories))/first(calories)*100, 4))

# makes a dataframe for the mean calorie intake for each food
meanveg <- veg_tidy %>% dplyr::group_by(food_group) %>% dplyr::summarise(Mean = round(mean(calories, na.rm = TRUE), 2))

# takes the top 25 foods that have the highest mean calorie intake for analysis and gets rid of foods that have na values for certain years
topveg <- head(meanveg %>% arrange(desc(Mean)), 20)
meannaveg <- veg_tidy %>% dplyr::group_by(food_group) %>% dplyr::summarise(Mean = round(mean(calories, na.rm = TRUE), 2))
topveg <- head(meannaveg %>% arrange(desc(Mean)), 20)

topveg_tidy <- subset(veg_tidy, food_group %in% topveg$food_group)
# bar graph of the top 25 foods that contribute the most to total caloric intake
topveg %>%
  mutate(food_group=factor(food_group, levels=food_group)) %>%
  ggplot()+
    geom_bar(aes(x=food_group, y=Mean), stat="identity") +
    ylab("Mean Daily Per Capita Calories")+
    xlab("Foods")+
    ggtitle("Top 20 Calorie Intake Vegetable Sources")+
    theme(axis.text.x = element_text(angle=60, hjust=1))

topveg_inc = merge(topveg_tidy, distinct(canc_cals_tidy[c("Year", "Rate")]), by="Year")


# Cancer incidence and top 25 food calories
ggplot(topveg_inc, aes(perc_dif, Rate))+
    geom_point()+
    geom_smooth(method=lm, color="blue", fill="blue")+
    ylab("Rate (per 100,000 people)")+
    xlab("Percent Difference Per Capita Calorie Intake from 2000")+
    xlim(-100, 100)+
    ggtitle("Cancer Incidence vs Calorie Intake per Top 20 Vegetables")+
    theme(axis.text.x = element_text(size=5.5, angle = 45))+
    facet_wrap(~food_group)


```

The clearest take-away here is that when the summary data say "vegetables" they really mean "potatoes." Everything else, by volume of caloric intake, has a much smaller effect, but is still interesting to see. And of course, potatoes are strongly associated with digestive cancer. Head lettuce and sweet corn are also strongly associated with digestive cancer incidence while sweet potatoes, broccoli and chili peppers are not.

```{r}
# top fruits and vegetables
fruit_inc = merge(fruits, sub_inc, by="Year")

fruit_inc_tidy <- gather(fruit_inc, foods, calories, colnames(fruit_inc)[!colnames(fruit_inc) %in% c("Year", "Sex", "Population", "Cases")], -c(Year, Rate))
fruit_inc_tidy <- veg_inc_tidy[complete.cases(fruit_inc_tidy),]

# converts the dataframe into a tidy format and changes the calorie values into numeric
fruit_tidy <- tidy_calorie_data(fruits)
fruit_tidy$calories <- as.numeric(fruit_tidy$calories)
fruit_tidy <- fruit_tidy %>% dplyr::group_by(food_group) %>% dplyr::filter(!any(is.na(calories)))

# gets the percent difference from 2000 for each year for each food
fruit_tidy <- fruit_tidy %>% dplyr::group_by(food_group) %>%
  dplyr::mutate(perc_dif = round((calories-first(calories))/first(calories)*100, 4))

# makes a dataframe for the mean calorie intake for each food
meanfruit <- fruit_tidy %>% dplyr::group_by(food_group) %>% dplyr::summarise(Mean = round(mean(calories, na.rm = TRUE), 2))

# takes the top 25 foods that have the highest mean calorie intake for analysis and gets rid of foods that have na values for certain years
topfruit <- head(meanfruit %>% arrange(desc(Mean)), 20)
meannafruit <- fruit_tidy %>% dplyr::group_by(food_group) %>% dplyr::summarise(Mean = round(mean(calories, na.rm = TRUE), 2))
topfruit <- head(meannafruit %>% arrange(desc(Mean)), 20)

topfruit_tidy <- subset(fruit_tidy, food_group %in% topfruit$food_group)
# bar graph of the top 25 foods that contribute the most to total caloric intake
topfruit %>%
  mutate(food_group=factor(food_group, levels=food_group)) %>%
  ggplot()+
    geom_bar(aes(x=food_group, y=Mean), stat="identity") +
    ylab("Mean Daily Per Capita Calories")+
    xlab("Foods")+
    ggtitle("Top 20 Calorie Intake Vegetable Sources")+
    theme(axis.text.x = element_text(angle=60, hjust=1))

topfruit_inc = merge(topfruit_tidy, distinct(canc_cals_tidy[c("Year", "Rate")]), by="Year")


# Cancer incidence and top 25 food calories
ggplot(topfruit_inc, aes(perc_dif, Rate))+
    geom_point()+
    geom_smooth(method=lm, color="blue", fill="blue")+
    ylab("Rate (per 100,000 people)")+
    xlab("Percent Difference Per Capita Calorie Intake from 2000")+
    xlim(-100, 100)+
    ggtitle("Cancer Incidence vs Calorie Intake per Top 20 Vegetables")+
    theme(axis.text.x = element_text(size=5.5, angle = 45))+
    facet_wrap(~food_group)


```

The same story can be told about fruits, but maybe not as well. Orange juice is strongly associated with digestive cancer while also making up a large portion of *calories by fruit*. Apple and pineapple juice, as well as a number of canned fruits, show similar results. In a sense, these food sources are smuggling added sugar into the fruits category in the summary data above. That doesn't let the rest of the fruits off the hook. Some are "good" (strawberries, avocados, cranberry juice) while others are not (peaches, pears).

### Cancer Mortality

These graphs are a bit of a sanity check. Our cancer incidence and cancer mortality data are very similar, and we would expect to see the same results if we had compared our calories data to mortality instead of incidence.

```{r}
# uses the combined sex since male and female cancer have the same trend
sub_mort <- read_csv('resources/processed_data/combined_mortality.csv')

# merges the major food groups and total calorie data set with the mortality data set
total_cals_mort = merge(total_cals, sub_mort, by="Year")

# Same process of cleaning and manipulation as the incidence
mort_cals_tidy <- gather(total_cals_mort, food_group, calories, colnames(total_cals_mort)[!colnames(total_cals_mort) %in% c("Year", "Sex", "Population", "Cases")], -c(Year, Rate))
mort_cals_tidy <- mort_cals_tidy[complete.cases(mort_cals_tidy),]

# gets percent difference
mort_cals_tidy <- mort_cals_tidy %>% dplyr::group_by(food_group) %>%
  dplyr::mutate(perc_dif = round((calories-first(calories))/first(calories)*100, 4))

#splits into food groups and total calories
mort_food_tidy <- subset(mort_cals_tidy, !food_group %in% c("Total", "Total (excluding Fats and Oils)"))

mort_total_tidy <- subset(mort_cals_tidy, food_group %in% c("Total", "Total (excluding Fats and Oils)"))


mort_cal_cor <- mort_cals_tidy %>% dplyr::group_by(food_group) %>% dplyr::summarise(Correlation = round(cor(perc_dif, Rate), 2))
temp <- mort_cals_tidy %>% dplyr::group_by(food_group) %>% dplyr::summarise(p_value = round(cor.test(perc_dif, Rate)$p.value, 4))
mort_cal_cor$p_value <- temp$p_value

# # Keep this graph (7)
cal_inc <- canc_cal_cor %>%
  arrange(desc(Correlation)) %>%
  mutate(food_group=factor(food_group, levels=food_group)) %>%
  ggplot(aes(x=food_group, y=Correlation))+
    geom_bar(stat="identity") +
    theme(axis.text.x = element_text(angle=60, hjust=1))

cal_mort <- mort_cal_cor %>%
  arrange(desc(Correlation)) %>%
  mutate(food_group=factor(food_group, levels=food_group)) %>%
  ggplot(aes(x=food_group, y=Correlation))+
    geom_bar(stat="identity") +
    theme(axis.text.x = element_text(angle=60, hjust=1))

#combines all foods with mortality dataframe then gets correlation between top foods and mortality for graphs
total_foods_mort = merge(topnafoods_tidy, distinct(mort_cals_tidy[c("Year", "Rate")]), by="Year")
food_mort_cor <- total_foods_mort %>% dplyr::group_by(food_group) %>% dplyr::summarise(Correlation = round(cor(perc_dif, Rate), 2))
temp <- total_foods_mort %>% dplyr::group_by(food_group) %>% dplyr::summarise(p_value = round(cor.test(perc_dif, Rate)$p.value, 4))
food_mort_cor$p_value <- temp$p_value

# Comparison between top 25 foods for cancer incidence and mortality
foods_inc <- food_inc_cor %>%
  arrange(desc(Correlation)) %>%
  mutate(food_group=factor(food_group, levels=food_group)) %>%
  ggplot(aes(x=food_group, y=Correlation))+
    geom_bar(stat="identity") +
    ggtitle("Incidence")+
    ylim(-1, 1)+
    xlab("Foods")+
    theme(axis.text.x = element_text(angle=60, hjust=1, size=8))

foods_mort <- food_mort_cor %>%
  arrange(desc(Correlation)) %>%
  mutate(food_group=factor(food_group, levels=food_group)) %>%
  ggplot(aes(x=food_group, y=Correlation))+
    geom_bar(stat="identity") +
    ggtitle("Mortality")+
    ylim(-1, 1)+
    xlab("Foods")+
    theme(axis.text.x = element_text(angle=60, hjust=1, size=8))

combined_foods <- foods_inc + foods_mort
combined_foods + plot_layout(guides = "collect") + plot_annotation("Top Food Correlations: Incidence vs Mortality")
combined_cals <- cal_mort + cal_inc
combined_cals + plot_layout(guides = "collect") + plot_annotation("Total Calorie/Major Food Group Intake Correlations: Incidence vs Mortality")
```

Albeit to different degrees, that is indeed the case.
