
```{r}
library(ggplot2)
library(tidyverse)
library(grid)
require(gridExtra)
require(scales)
library(dplyr)
library(patchwork)

```

```{r include=FALSE}
dairy <- read_csv("resources/processed_data/dairy.csv")
fats <- read_csv("resources/processed_data/fats.csv")
fruits <- read_csv("resources/processed_data/fruits.csv")
grains <- read_csv("resources/processed_data/grains.csv")
incidence <- read_csv("resources/processed_data/incidence.csv")
meateggsnuts <- read_csv("resources/processed_data/meateggsnuts.csv")
mortality <- read_csv("resources/processed_data/mortality.csv")
sugars <- read_csv("resources/processed_data/sugars.csv")
vegetables <- read_csv("resources/processed_data/vegetables.csv")
total_cals <- read_csv("resources/processed_data/total.csv")

tidy_calorie_data <- function(food_table) {
  df <- subset(gather(food_table, food_group, calories,    colnames(food_table), -Year), Year > 1999)
  return(df)
}
```

## Results

1. Line graph of mortality and incidence
2. Line graph of total caloric intake and caloric intake by major food group
3. Line graph of specific foods
4. Scatterplot of total caloric intake and caloric intake by major food group vs cancer mortality
5. Scatterplot of total caloric intake and caloric intake by major food group vs cancer incidence
6. Scatterplot of specific foods vs cancer incidence/mortality
7. Bargraph of R-squared values/correlation coefficients
8. Percent difference graphs with incidence and caloric intake

## Digestive Cancer Trends

```{r include=FALSE}

# generates the cancer incidence plots
inc <-ggplot(data=incidence)+
  geom_line(aes(x=Year, y=Rate, color=Sex)) +
  geom_point(aes(x=Year, y=Rate, color=Sex))+
  scale_color_manual(values=c("black", "red", "blue"))+
  theme(legend.position = "none")+
  ggtitle("Digestive System Cancer Incidence Rates")+
  ylab("Rate (per 100,000 people)")
  
# generates the cancer mortality plots
mort <- ggplot(data=mortality)+
  geom_line(aes(x=Year, y=Rate, color=Sex)) +
  geom_point(aes(x=Year, y=Rate, color=Sex))+
  ggtitle("Digestive System Cancer Mortality Rates")+
  ylab("Rate (per 100,000 people)")+
  scale_color_manual(values=c("black", "red", "blue"))

#fix axes
# takes the incidence and mortality plots and puts them on the same line with shared legend (1)
combined <- inc + mort & theme(legend.position = "bottom")
combined + plot_layout(guides = "collect")
```

These figures show the trends of digestive system cancer incidence and mortality over the years. As we see here, both the digestive system incidence and mortality rates have been decreasing since 2000, with mortality sharply decreasing after 2014/2015. These graphs also show that men are at a greater risk of developing digestive cancer compared to women and the trends of incidence and mortality for both genders have similar shapes. Now that we have established the trends of digestive system cancer, let's explore dietary trends from 2000 to 2017

## Dietary Trends

### Food Group and Overall Calorie Consumption Trends 

```{r}
# sums all the columns except for the fats and oils column to get the total amount of calories consumed besides the fats and oils
total_cals$`Total (excluding Fats and Oils)` <- rowSums(total_cals[,c(2:6, 8)])

# creates two separate dfs: one for the food groups (tidy_total) and one for the totals (tidy_all)
tidy_total = tidy_calorie_data(total_cals[-c(9,10)])
tidy_all = tidy_calorie_data(total_cals[c(1,9:10)])

cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
# graph of the per capita total calorie intake and total calorie intake excluding fats and oils
ggplot(data=tidy_all, aes(x=Year, y=calories, color=food_group)) +
  geom_line() +
  geom_point()+
  ggtitle("Total Daily Per Capita Calorie Intake from 2000-2017")

# graph of the per capita calorie intake by major food groups
ggplot(data=tidy_total, aes(x=Year, y=calories, color=food_group)) +
  scale_fill_manual(values=cbPalette)+
  geom_line() +
  geom_point()+
  ylab("Daily Per Capita Calories")+
  ggtitle("Daily Per Capita Calorie Intake by Major Food Group from 2000-2017")
```
These graphs show the trend of both the per capita total daily calorie intake and the calorie intake divided by major food groups. The most notable changes for the food groups were sugar and sweeteners, flour and cereal products, and meat, eggs, and nuts. Both sugar and sweeteners and flour and ceral products decreased from 2000 to 2017 while meat, eggs, and nuts increased from 2014-2017 after gradually decreasing before. Intake of fruits and vegetables has also slightly decreased over the years. In terms of calories, both total calories and total calories, excluding fats and oils have been decreasing from 2000-2017 as a whole. However, from 2011 to 2017 total calories excluding fats and oils has risen slightly. Calorie intake data for fats and oils from 2010 onward could not be found leading to the creation of the total calorie intake excluding fats and oils variable. Now that we know that calorie intake has been decreasing as a whole and within many major food groups, let's look at certain foods within each food group to discover more trends.

### Select Food Trends

```{r}
# merges all the separate food group dataframes into one dataframe joined by year
allfoods = merge(meateggsnuts[-c(1)], 
                 merge(sugars[-c(1)], 
                 merge(dairy[-c(1)], 
                 merge(fats[-c(1)], 
                 merge(fruits[-c(1)], merge(grains[-c(1)], vegetables[-c(1)], by ="Year"), by ="Year"), by ="Year"), by ="Year"), by ="Year"), by ="Year")

# converts the dataframe into a tidy format and changes the calorie values into numeric
allfoods_tidy <- tidy_calorie_data(allfoods)
allfoods_tidy$calories <- as.numeric(allfoods_tidy$calories)
allfoods_na <- allfoods_tidy %>% dplyr::group_by(food_group) %>% dplyr::filter(!any(is.na(calories)))

# gets the percent difference from 2000 for each year for each food
allfoods_tidy <- allfoods_tidy %>% dplyr::group_by(food_group) %>%
  dplyr::mutate(perc_dif = round((calories-first(calories))/first(calories)*100, 4)) 

# makes a dataframe for the mean calorie intake for each food
meanfoods <- allfoods_tidy %>% dplyr::group_by(food_group) %>% dplyr::summarise(Mean = round(mean(calories, na.rm = TRUE), 2))

# takes the top 25 foods that have the highest mean calorie intake for analysis and gets rid of foods that have na values for certain years
topfoods <- head(meanfoods %>% arrange(desc(Mean)), 25)
meannafoods <- allfoods_na %>% dplyr::group_by(food_group) %>% dplyr::summarise(Mean = round(mean(calories, na.rm = TRUE), 2))
topnafoods <- head(meannafoods %>% arrange(desc(Mean)), 25)

# bar graph of the top 25 foods that contribute the most to total caloric intake
topnafoods %>%
  mutate(food_group=factor(food_group, levels=food_group)) %>%
  ggplot()+
    geom_bar(aes(x=food_group, y=Mean), stat="identity") +
    ylab("Mean Daily Per Capita Calories")+
    xlab("Foods")+
    ggtitle("Top 25 Calorie Intake Food Sources")+
    theme(axis.text.x = element_text(angle=60, hjust=1))
  
# splits the 25 foods into low medium and high to increase clarity of the line plot
topnafoods_tidy <- subset(allfoods_tidy, food_group %in% topnafoods$food_group)
topnafoods_tidy$calories_consumed <- cut(topnafoods_tidy$calories, breaks = c(0,50,100,500), label=c("low", "medium", "high"))

# line plot of the top 25 foods
ggplot(data=topnafoods_tidy) +
  geom_line(aes(x=Year, y=calories, color=food_group)) +
  ggtitle("Daily Per Capita Calorie Intake for Top 25 Food Sources from 2000-2017")+
  ylab("Mean Daily Per Capita Calories")+
  xlab("Years")+
  facet_wrap(~calories_consumed, scales = "free") +
  theme(axis.text.x = element_text(angle=60, hjust=1))+
  theme(legend.position = "right")

```
In order to determine how diet is associated with digestive cancer incidence and mortality, we took the top 25 foods that contributed the most to daily caloric intake and analyzed their changes from 2000 to 2017. These food sources included several food sources that are related to digestive cancer such as beef and high fructose corn syrup. Other foods that surprising cracked the top 25 were bananas, cane/beet sugar, and  frozen potatoes. The highest caloric contributer was white and whole flour by a large margin. As shown in the line graph, these food groups changed dynamically from 2000 to 2017 as a whole with foods such as peanuts, chicken, and butter increasing and other foods such as beef, high fructose corn syrup, and ice cream decreasing. Now that we've set the stage, let's jump into the comparisons between cancer incidence/mortality and foods/total calorie intake.


## Comparison of Food/Caloric Intake and Digestive Cancer Trends

### Cancer Incidence
```{r}
# uses the combined sex since male and female cancer have the same trend
sub_inc <- subset(incidence, Sex == "Combined")

# used to combine the incidence and major food group/total calories dataframes then tidy
total_cals_inc = merge(total_cals, sub_inc, by="Year")
canc_cals_tidy <- gather(total_cals_inc, food_group, calories, colnames(total_cals_inc)[!colnames(total_cals_inc) %in% c("Year", "Sex", "Population", "Cases")], -c(Year, Rate))
canc_cals_tidy <- canc_cals_tidy[complete.cases(canc_cals_tidy),]

# Splits the dataframe into the major food groups and the totals
canc_food_tidy <- subset(canc_cals_tidy, !food_group %in% c("Total", "Total (excluding Fats and Oils)"))
canc_total_tidy <- subset(canc_cals_tidy, food_group %in% c("Total", "Total (excluding Fats and Oils)"))

# Graph between the calories
ggplot(canc_food_tidy, aes(calories, Rate))+
    geom_point()+
    geom_smooth(method=lm, color="blue", fill="blue")+
    ylab("Rate (per 100,000 people)")+
    xlab("Daily Per Capita Calorie Intake")+
    ggtitle("Cancer Incidence vs Calorie Intake per Major Food Group")+
    facet_wrap(~food_group)
```
Even with using a smoother to draw a line through thte data, the correlations are somewhat unclear using the raw caloric intake values. We can clearly see that both sugar and sweeteners and flour and cereal products correlate with digestive cancer incidence but the trends of the other groups is unclear. We will use percent difference from 2000 per capita caloric intake from now on to show the trends more clearly.

```{r}
# Gets the percent difference from 2000
canc_cals_tidy <- canc_cals_tidy %>% dplyr::group_by(food_group) %>%
  dplyr::mutate(perc_dif = round((calories-first(calories))/first(calories)*100, 4)) 

# Creates a dataframe that has the correlation coefficient between the food group/total and cancer incidence and p_value for correlation
canc_cal_cor <- canc_cals_tidy %>% dplyr::group_by(food_group) %>% dplyr::summarise(Correlation = round(cor(perc_dif, Rate), 2))
temp1 <- canc_cals_tidy %>% dplyr::group_by(food_group) %>% dplyr::summarise(p_value = round(cor.test(perc_dif, Rate)$p.value, 4))
canc_cal_cor$p_value <- temp1$p_value

# Cancer Incidence vs total caloric intake and major food groups
ggplot(canc_cals_tidy, aes(perc_dif, Rate))+
    geom_point()+
    geom_smooth(method=lm, color="blue", fill="blue")+
    ylab("Rate (per 100,000 people)")+
    xlab("Percent Difference from 2000 Per Capita Calorie Intake")+
    ggtitle("Cancer Incidence vs Calorie Intake per Major Food Group/Total Calorie Intake")+
    facet_wrap(~food_group)

# Correlation coefficients for above graph
canc_cal_cor %>% 
  arrange(desc(Correlation)) %>%
  mutate(food_group=factor(food_group, levels=food_group)) %>%
  ggplot(aes(x=food_group, y=Correlation))+
    geom_bar(stat="identity") +
    geom_text(aes(label=p_value), size=2, color="red", position = position_stack(vjust = .5))+
    annotate("text", label = "Red text within bar graph is the p value associated with correlation", x = 5, y = -1, size = 2, colour = "red")+
    ylab("Correlation Coefficient")+
    xlab("Food Groups")+
    ggtitle("Correlation Coefficients of Food Groups/Total Caloric Intake vs Cancer Incidence")+
    theme(axis.text.x = element_text(angle=60, hjust=1))


total_foods_inc = merge(topnafoods_tidy, distinct(canc_cals_tidy[c("Year", "Rate")]), by="Year")


food_inc_cor <- total_foods_inc %>% dplyr::group_by(food_group) %>% dplyr::summarise(Correlation = round(cor(perc_dif, Rate), 2))
temp <- total_foods_inc %>% dplyr::group_by(food_group) %>% dplyr::summarise(p_value = round(cor.test(perc_dif, Rate)$p.value, 4))
food_inc_cor$p_value <- temp$p_value

# Cancer incidence and top 25 food calories
ggplot(total_foods_inc, aes(perc_dif, Rate))+
    geom_point()+
    geom_smooth(method=lm, color="blue", fill="blue")+
    ylab("Rate (per 100,000 people)")+
    xlab("Percent Difference from 2000 Per Capita Calorie Intake")+
    ggtitle("Cancer Incidence vs Calorie Intake per Top 25 Foods")+
    facet_wrap(~food_group)


# Correlation coefficients for above graph
food_inc_cor %>% 
  arrange(desc(Correlation)) %>%
  mutate(food_group=factor(food_group, levels=food_group)) %>%
  ggplot(aes(x=food_group, y=Correlation))+
    geom_bar(stat="identity") +
    geom_text(aes(label=p_value), size=2, color="red", position = position_stack(vjust = .5))+
    annotate("text", label = "Red text within bar graph is the p value associated with correlation", x = 5, y = -1, size = 2, colour = "red")+
    ylab("Correlation Coefficient")+
    xlab("Foods")+
    ggtitle("Correlation Coefficients of Top 25 Foods vs Cancer Incidence")+
    theme(axis.text.x = element_text(angle=60, hjust=1))


# total_cals_inc$`Sugar and sweeteners (Added)`
# ggplot(total_cals_inc) +
#   geom_point(aes(`Sugar and sweeteners (Added)`, Rate, color=Sex))
# 
# sugar_inc = merge(sugars, sub_inc, by="Year")
# sugar_inc$`Cane and Beet Sugar`
# ggplot(sugar_inc) +
#   geom_point(aes(`Cane and Beet Sugar`, Rate, color=Sex))
# 
# meat_inc = merge(meateggsnuts, sub_inc, by="Year")
# ggplot(meat_inc) +
#   geom_point(aes(Peanuts, Rate, color=Sex))
# 
# cor(meat_inc$Beef, test$Rate)
# cor.test(meat_inc$Beef, test$Rate)

```
We found that the decrease in total calorie intake correlated with the decrease in digestive cancer incidence. We also found that food groups such as sweeteners and sugars and cereal/flour products positively correlated with digestive cancer incidence. Surprisingly, we found that fruits caloric intake and vegetable caloric intake correlated positively with digestive cancer incidence. We think that since total caloric intake is associated with cancer incidence, the other food groups appeared to be correlated as well. For the top 25 foods however, we had some more interesting results. As expected, we found that high fructose corn syrup and beef were positively associated with digestive cancer incidence. We also saw ice cream, potatoes, and orange juice trending positively with incidence. On the other hand, we found that butter, peanuts, cane/beet sugar, and mozzarella trended negatively with cancer incidence, meaning the higher the calorie intake of these foods, the lower the incidence.


### Cancer Mortality
```{r}
# uses the combined sex since male and female cancer have the same trend
sub_mort <- subset(mortality, Sex == "Combined")
# merges the major food groups and total calorie data set with the mortality data set
total_cals_mort = merge(total_cals, sub_mort, by="Year")
# Same process of cleaning and manipulation as the incidence
mort_cals_tidy <- gather(total_cals_mort, food_group, calories, colnames(total_cals_mort)[!colnames(total_cals_mort) %in% c("Year", "Sex", "Population", "Cases")], -c(Year, Rate))
mort_cals_tidy <- mort_cals_tidy[complete.cases(mort_cals_tidy),]
# gets percent difference
mort_cals_tidy <- mort_cals_tidy %>% dplyr::group_by(food_group) %>%
  dplyr::mutate(perc_dif = round((calories-first(calories))/first(calories)*100, 4)) 

#splits into food groups and total calories
mort_food_tidy <- subset(mort_cals_tidy, !food_group %in% c("Total", "Total (excluding Fats and Oils)"))

mort_total_tidy <- subset(mort_cals_tidy, food_group %in% c("Total", "Total (excluding Fats and Oils)"))


# mort_cal_cor <- mort_cals_tidy %>% dplyr::group_by(food_group) %>% dplyr::summarise(Correlation = round(cor(perc_dif, Rate), 2))
# temp <- mort_cals_tidy %>% dplyr::group_by(food_group) %>% dplyr::summarise(p_value = round(cor.test(perc_dif, Rate)$p.value, 4))
# mort_cal_cor$p_value <- temp$p_value


# # Keep this graph (5)
# ggplot(mort_food_tidy, aes(perc_dif, Rate))+
#     geom_point()+
#     geom_smooth(method=lm, color="blue", fill="blue")+
#     facet_wrap(~food_group)
# 
# # Keep this graph (6)
# ggplot(mort_total_tidy, aes(perc_dif, Rate))+
#     geom_point()+
#     geom_smooth(method=lm, color="blue", fill="blue")+
#     facet_wrap(~food_group)
# 
# # Keep this graph (7)
# mort_cal_cor %>% 
#   arrange(desc(Correlation)) %>%
#   mutate(food_group=factor(food_group, levels=food_group)) %>%
#   ggplot(aes(x=food_group, y=Correlation))+
#     geom_bar(stat="identity") +
#     geom_text(aes(label=p_value), size=2, color="red", position = position_stack(vjust = .5))+
#     annotate("text", label = "Red text within bar graph is the p value associated with correlation", x = 5, y = -1, size = 2, colour = "red")+
#     theme(axis.text.x = element_text(angle=60, hjust=1))

#combines all foods with mortality dataframe then gets correlation between top foods and mortality for graphs
total_foods_mort = merge(topnafoods_tidy, distinct(mort_cals_tidy[c("Year", "Rate")]), by="Year")
food_mort_cor <- total_foods_mort %>% dplyr::group_by(food_group) %>% dplyr::summarise(Correlation = round(cor(perc_dif, Rate), 2))
temp <- total_foods_mort %>% dplyr::group_by(food_group) %>% dplyr::summarise(p_value = round(cor.test(perc_dif, Rate)$p.value, 4))
food_mort_cor$p_value <- temp$p_value

# # Keep this graph (5)
# ggplot(total_foods_mort, aes(perc_dif, Rate))+
#     geom_point()+
#     geom_smooth(method=lm, color="blue", fill="blue")+
#     facet_wrap(~food_group)
# 
# 
# # Keep this graph (7)
# food_mort_cor %>% 
#   arrange(desc(Correlation)) %>%
#   mutate(food_group=factor(food_group, levels=food_group)) %>%
#   ggplot(aes(x=food_group, y=Correlation))+
#     geom_bar(stat="identity") +
#     geom_text(aes(label=p_value), size=2, color="red", position = position_stack(vjust = .5))+
#     annotate("text", label = "Red text within bar graph is the p value associated with correlation", x = 5, y = -1, size = 2, colour = "red")+
#     theme(axis.text.x = element_text(angle=60, hjust=1))


# Comparison between top 25 foods for cancer incidence and mortality
foods_inc <- food_inc_cor %>% 
  arrange(desc(Correlation)) %>%
  mutate(food_group=factor(food_group, levels=food_group)) %>%
  ggplot(aes(x=food_group, y=Correlation))+
    geom_bar(stat="identity") +
    ggtitle("Incidence")+
    theme(axis.text.x = element_text(angle=60, hjust=1))

foods_mort <- food_mort_cor %>% 
  arrange(desc(Correlation)) %>%
  mutate(food_group=factor(food_group, levels=food_group)) %>%
  ggplot(aes(x=food_group, y=Correlation))+
    geom_bar(stat="identity") +
    ggtitle("Mortality")+
    theme(axis.text.x = element_text(angle=60, hjust=1))

combined_foods <- foods_inc + foods_mort
combined_foods + plot_layout(guides = "collect") + plot_annotation("Incidence Correlations vs Mortality Correlations")
```
Through this graph we can see that the same foods that were correlated with the increase/decrease in cancer incidence were also associated with the increase/decrease of cancer mortality, albiet to slightly different degrees.
